---
title: "Power Analysis for Final Project"
author: "Cynthia, Emanuel, Jonathan, and Rina "
date: "02/14/2024"
name: "Rina Palta"
output: 
  pdf_document: 
    number_sections: true
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(include = FALSE, message = FALSE, warning = FALSE )

knitr::knit_engines$set(problem_description = function(options) {
  code <- paste(options$code, collapse = "\n")
})
```

```{r package loads, warning=FALSE, message=FALSE}
library(data.table)
library(sandwich)
library(lmtest)

library(ggplot2)
library(knitr)
```

Sleep score ranges from 0-100, with most sleep scores falling in the 60-85 range. We could do a normally distributed variable centered at 72, with a standard error of 6. And if we're looking for an ATE of 3, for the treatment group, we can shift that distribution up by 3, keeping the distribution for the control group the same. 


```{r}
d <- data.table()

d[ , condition := rep(c('control', 'treatment'), each = 100)]
d[condition == 'control', Y := rnorm(.N, mean = 72, sd = 6)]
d[condition == 'treatment', Y := rnorm(.N, mean = 75, sd = 6)]

d[ , t.test(Y ~ condition)]
```
```{r}
d
```


The true ATE in our 200-person sample is 1.89, with a 95% confidence interval of .29 to .3.48. 

Simulating the experiment at different sample sizes, from 10 - 200% of our datatable will give us an idea of what the power of our experiment will be at different sample sizes. 

```{r}

# vector of percentages to sample
percentages_to_sample <- seq(0.10, 2.00, by = 0.10)

# empty vector to store the power calculations at every sample size
power_results <- rep(NA, length(percentages_to_sample))

# number of observations in the original dataset
n_obs <- nrow(d)

# loop through percentages to sample
for (j in seq_along(percentages_to_sample)) {
  # calculate the sample size for each group, control and treatment
  sample_size <- round(percentages_to_sample[j] * n_obs) / 2
  
  # empty vector to store the p-values at the current sample size
  t_test_p_values <- rep(NA, 1000) 
  
  # run the experiment at the current sample size 1000 times and store the 1000 p-values created from each t-test
  for (i in 1:1000) {
    # sample from treatment and control--half in treatment and half in control
    sample_treatment <- d[condition == 'treatment'][sample(.N, sample_size, replace = TRUE)]
    sample_control <- d[condition == 'control'][sample(.N, sample_size, replace = TRUE)]
    
    # combine treatment and control groups
    next_t_test <- rbindlist(list(sample_treatment, sample_control))
    
    # perform t-test
    next_test_result <- next_t_test[ , t.test(Y ~ condition)]
    
    # store the p-value
    t_test_p_values[i] <- next_test_result$p.value
  }
  
  # Calculate power and store in power_results
  power_results[j] <- mean(t_test_p_values <= 0.05)
}


```

```{r}
plot(power_results_table$sample_size, power_results_table$test_power)

```

Note that the sample size will actually need to be half of what appears in the chart to reach the given power, since we have each subject participating in control and treatment. 
