---
title: "Power Analysis for Final Project"
author: "Cynthia, Emanuel, Jonathan, and Rina "
date: "02/14/2024"
name: "Rina Palta"
output: 
  pdf_document: 
    number_sections: true
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(include = FALSE, message = FALSE, warning = FALSE )

knitr::knit_engines$set(problem_description = function(options) {
  code <- paste(options$code, collapse = "\n")
})
```

```{r package loads, message=FALSE, warning=FALSE}
library(data.table)
library(sandwich)
library(lmtest)

library(ggplot2)
library(knitr)
```

# Introduction

Our main research question is: Does wearing sunglasses before bedtime improve sleep quality? Where sleep quality is measured via wearable technology, such as a smart watch. Sleep quality itself can be operationalized in various ways. We will be looking specifically at sleep quality as measured by

1.  Sleep Score. This is a proprietary metric whose calculation may vary from app to app but we are taking these differences as approximately the same.
2.  Duration spent in REM and deep sleep. Time spent in specific stages of sleep, particularly REM and deep sleep, have been found to be essential for high quality sleep. (\textit{Stages of sleep: What happens in a sleep cycle})

Our assumptions for each power scenario are:

-   Population is normally distribution.
-   Significance level = 0.05.

We simulate the experiment at different sample sizes, from 10-100 subjects by increments of 10. We run 1,000 simulations at each sample size to calculate the power. We use t-test for hypothesis testing. Note: Because we are running a paired test, each subject has a control and treatment value; for example, 10 subjects yields 10 control and 10 treatment sleep measurements.

```{r include=FALSE}
# Assumptions
control_avg_score = 75

## Other parameters

#significance level
alpha = 0.05

# vector of sample sizes
sample_sizes = seq(10, 100, by = 10)
```

# Scenario 1 - Measurement: Sleep Score, ATE: 5, Variance: 5

For this scenario we assume:

-   The average non-intervention sleep score is 75 (for scenarios 1, 2, & 3). This average is based off of data provided by Garmin (\textit{Garmin Blog}).
-   True ATE = 5. This is based on a small trial run of our experiment conducted with ourselves.
-   Standard deviation = 5, regardless of intervention

```{r echo=FALSE, include=TRUE}
jon_control_scores = c(70, 78, 72)
jon_treatment_scores = c(79, 85, 80)


cat('Average sleep score in control phase: ', mean(jon_control_scores))
cat('\nAverage sleep score in treatment phase: ', mean(jon_treatment_scores))

cat('\nJon pilot average treatment effect: ', mean(jon_treatment_scores) - mean(jon_control_scores))
```

```{r echo=FALSE, include=TRUE}
# Scenario parameters
true_ate = 5
stdev = 5


sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, proportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
ate_5 <- data.table(sample_sizes, sample_powers)
ate_5

#scenario power
power_1 <- ate_5[sample_sizes == 20, sample_powers]
```

In this scenario, we are able to achieve a power of `r ate_5[sample_sizes == 20, sample_powers]` with a sample size of 20.

# Scenario 2 - Measurement: Sleep Score, ATE: 3, Variance: 3

For this scenario we assume:

-   The average non-intervention sleep score is 75.
-   True ATE = 3
-   Standard deviation = 3, regardless of intervention

```{r echo=FALSE, include=TRUE}
# Scenario parameters
true_ate = 3
stdev = 3

sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]    
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, proportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
ate_3 <- data.table(sample_sizes, sample_powers)
ate_3
```

In this scenario, we are able to achieve a power of `r ate_3[sample_sizes == 20, sample_powers]` with a sample size of 20.

# Scenario 3 - Measurement: Sleep Score, ATE: 5, Variance: 7

For this scenario we assume:

-   The average non-intervention sleep score is 75.
-   True ATE = 5
-   Standard deviation = 7, regardless of intervention

```{r echo=FALSE, include=TRUE}
# Scenario parameters
true_ate = 5
stdev = 7

sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]    
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, proportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
ate_5_7 <- data.table(sample_sizes, sample_powers)
ate_5_7
```

# Scenario 4 - Measurement: Deep & REM sleep

We are interest in potentially looking at other sleep metrics as measurement of our outcome of sleep quality. We know that sleep quality, and not just duration, particularly the amount of time spent in the deep and REM phases of sleep can be strong indicators of sleep quality (\textit{Suni, E.}). We attempted to used the effect size calculated by Hedges' g in the study by \textit{Chinoy, E, et al.} in which they compared several different consumer sleep tracking devices to the laboratory standard of polysomnography.

For this scenario we use values from the  \textit{Chinoy, E, et al.} study

-   The average non-intervention deep sleep phase as a percentage of total sleep is approximately 15%.
-   The average non-intervention REM sleep phase as a percentage of total sleep is approximately 26%.
-   The true ATE of deep sleep is 0.185.
-   The true ATE of REM sleep is 0.25.
-   Standard deviation of deep sleep = 6%, regardless of intervention.
-   Standard deviation of REM sleep = 8%, regardless of intervention.

```{r eval=FALSE, include=FALSE}
# Define the mean and standard deviation of deep sleep and total sleep time
mean_deep_sleep <- 69.4
sd_deep_sleep <- 27.3
mean_total_sleep <- 456.8
sd_total_sleep <- 21.0

# Calculate the percentage of deep sleep relative to total sleep time
# Deep sleep percentage = (Deep sleep / Total sleep) * 100
deep_sleep_percentage <- (mean_deep_sleep / mean_total_sleep) * 100

# Use the formula for propagating errors to calculate the standard deviation of the percentage
# Reference: https://en.wikipedia.org/wiki/Propagation_of_uncertainty#Example_formulas
sd_deep_sleep_percentage <- deep_sleep_percentage * sqrt((sd_deep_sleep/mean_deep_sleep)^2 + (sd_total_sleep/mean_total_sleep)^2)

# Print the result
cat("Average deep sleep as a percentage of total sleep time:", round(deep_sleep_percentage, 2), "%\n")
cat("Standard deviation of deep sleep percentage:", round(sd_deep_sleep_percentage, 2), "%\n")
```

```{r eval=FALSE, include=FALSE}
# Define the mean and standard deviation of deep sleep and total sleep time
mean_rem_sleep <- 120
sd_rem_sleep <- 37.8
mean_total_sleep <- 456.8
sd_total_sleep <- 21.0

# Calculate the percentage of deep sleep relative to total sleep time
# Deep sleep percentage = (Deep sleep / Total sleep) * 100
rem_sleep_percentage <- (mean_rem_sleep / mean_total_sleep) * 100

# Use the formula for propagating errors to calculate the standard deviation of the percentage
# Reference: https://en.wikipedia.org/wiki/Propagation_of_uncertainty#Example_formulas
sd_rem_sleep_percentage <- rem_sleep_percentage * sqrt((sd_rem_sleep/mean_rem_sleep)^2 + (sd_total_sleep/mean_total_sleep)^2)

# Print the result
cat("Average REM sleep as a percentage of total sleep time:", round(rem_sleep_percentage, 2), "%\n")
cat("Standard deviation of REM sleep percentage:", round(sd_rem_sleep_percentage, 2), "%\n")

```

```{r echo=FALSE, include=TRUE}
# Scenario parameters
true_ate = 0.185
stdev = 6

sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]    
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, proportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
ate_deep <- data.table(sample_sizes, sample_powers)
ate_deep
```

```{r echo=FALSE, include=TRUE}
# Scenario parameters
true_ate = 0.25
stdev = 8

sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]    
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, proportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
ate_rem <- data.table(sample_sizes, sample_powers)
ate_rem
```
We had some difficulty in interpreting the units of the Hedges' g effect size and how it factors into our power calculation. From these power results, we can see we have extremely low statistical power. We also looked at the deep & REM phases of our self-conducted trial study and found that the difference in means between treatment and control for time spent in deep & REM phases of sleep was also extremely small (not even one-tenth of a percentage point). It is therefore unlikely we can feasibly use these metrics for our outcome.

# Plot of Achieved Power

Comparing the powers of each scenario, we plot:

```{r echo=FALSE, include=TRUE}
#list of data.tables
data_tables <- list(ate_5, ate_3, ate_5_7, ate_rem)

combined_data <- do.call(rbind, data_tables)
combined_data$dataset <- rep(1:length(data_tables), sapply(data_tables, nrow))

# Plot combined data with grouping
ggplot(combined_data, aes(x = sample_sizes, y = sample_powers, group = dataset, color = as.factor(dataset))) +
  geom_smooth(method = "loess") +
  geom_point() +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "red") + 
  labs(x = "Sample Size", y = "Power", title = "Power Analysis", color = "Scenario") +
  theme(legend.title = element_text(face="bold"))
```
We reach a power of .8 around 26-27 subjects. 

```{=tex}
\section{References}
\begin{small}
\begin{itemize}
\item[] Evan D Chinoy, Joseph A Cuellar, Kirbie E Huwa, Jason T Jameson, Catherine H Watson, Sara C Bessman, Dale A Hirsch, Adam D Cooper, Sean P A Drummond, Rachel R Markwald, Performance of seven consumer sleep-tracking devices compared with polysomnography, Sleep, Volume 44, Issue 5, May 2021, zsaa291, \url{https://doi.org/10.1093/sleep/zsaa291}

\item[] Miller DJ, Sargent C, Roach GD. A Validation of Six Wearable Devices for Estimating Sleep, Heart Rate and Heart Rate Variability in Healthy Adults. Sensors. 2022; 22(16):6317.
\url{https://doi.org/10.3390/s22166317}

\item[] Suni, E., Singh, A. (2023, December 8). Stages of sleep: What happens in a sleep cycle. Sleep Foundation. \url{https://www.sleepfoundation.org/stages-of-sleep}

\item[] New data examines quality of Garmin users’ sleep. Garmin Blog. (2023, August 30). \url{https://www.garmin.com/en-US/blog/health/new-data-examines-quality-of-garmin-users-sleep/}

\end{itemize}
\end{small}
```


