---
title: "Power Analysis for Final Project"
author: "Cynthia, Emanuel, Jonathan, and Rina "
output: 
  pdf_document: 
    number_sections: true
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(include = FALSE, message = FALSE, warning = FALSE )

knitr::knit_engines$set(problem_description = function(options) {
  code <- paste(options$code, collapse = "\n")
})
```

```{r package loads, warning=FALSE, message=FALSE}
library(data.table)
library(sandwich)
library(lmtest)

library(ggplot2)
library(knitr)
```

My scores

-   No sunglasses: Wed 2/7 - Fri 2/9

-   Sunglasses: Wed 2/14 - Fri 2/16

```{r}
jon_control_scores = c(70, 78, 72)
jon_treatment_scores = c(79, 85, 80)

mean(jon_control_scores)
mean(jon_treatment_scores)

cat('Jon pilot average treatment effect: ', mean(jon_treatment_scores) - mean(jon_control_scores))

```

**Assume for population for each scenario**

-   Normal distribution

-   Average non-intervention sleep score 75

We simulate the experiment at different sample sizes, from 10-100 subjects by increments of 10. We run 1,000 simulations at each sample size to calculate the power. Note: Because we are running a paired test, each subject has a control and treatment value; for example, 10 subjects yields 10 control and 10 treatment sleep scores.

```{r}
# Assumptions
control_avg_score = 75

## Other parameters

#significance level
alpha = 0.05

# vector of sample sizes
sample_sizes = seq(10, 100, by = 10)
```

## **Scenario #1: Base Case**

-   True ATE = 5

-   Standard deviation = 5, regardless of intervention

```{r}
# Scenario parameters
true_ate = 5
stdev = 5


sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, propportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
data.table(sample_sizes, sample_powers)
```

## **Scenario #1A**

Same ATE, but greater standard deviation

-   True ATE = 5

-   Standard deviation = 10, regardless of intervention

```{r}
# Scenario parameters
true_ate = 5
stdev = 10

sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, propportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
data.table(sample_sizes, sample_powers)
```

## **Scenario #2**

Smaller ATE

-   True ATE = 3

-   Standard deviation = 3, regardless of intervention

```{r}
# Scenario parameters
true_ate = 3
stdev = 3

sample_powers = NA   # Create vector to store power at different sample sizes

for (i in 1:length(sample_sizes)) {
  
  sample_size_p_values = NA # Create vector to store p-values

  for (sim in 1:1000) {
    d <- data.table() # Create empty data table
    
    # Draw from control and treatment population
    # Paired test, treatment effect is added to control
    d[, control:= rnorm(sample_sizes[i], mean = control_avg_score, sd = stdev)]
    d[, treatment:= control + rnorm(sample_sizes[i], mean = true_ate, sd = stdev)]
    # Ensure score is between 0 and 100 inclusive 
    d[, control:= pmin(pmax(control, 0), 100)]
    d[, treatment:= pmin(pmax(treatment, 0), 100)]    
    
    # t-test of treatment compared control scores and get p-value
    # Store p-value in vector
    sample_size_p_values[sim] = t.test(d[, treatment], d[, control], conf.level=1-alpha)$p.value
  }
  
  # Calculate power, propportion of p_values below our alpha
  # Store in sample_power vector
  sample_powers[i] = mean(sample_size_p_values < alpha)

}

# Show sample sizes and powers
data.table(sample_sizes, sample_powers)
```

**PLOT**
