---
title: "jon_EDA"
output: html_document
date: "2024-04-04"
---

```{r setup, message=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(lmtest)
library(sandwich)
```

```{r read data}
data <- fread("../data/paired_cleaned_data.csv")
data_treat <- fread("../data/cleaned_data_w_assign.csv")
data_long <- fread("../data/cleaned_data_long.csv")
```

```{r}


```

Relationship between subjective measures and sleep scores?

```{r}
# Average of sleep scrores for each subject-phase 
data_treat[, sleep_score_mean := rowMeans(.SD, na.rm = TRUE), .SDcols = c("day1_ss", "day2_ss", "day3_ss", "day4_ss")]
```

# Association between Sleep Score and Subjective Measures

```{r}
summary(data_treat[, lm(sleep_score_mean ~ phase_sq)])
```

```{r}
summary(data_treat[, lm(sleep_score_mean ~ phase_ts)])
```

Stat sig evaluations of these relationships supports the use of sleep score measures for outcome measure.

```{r}
summary(data_treat[, lm(phase_sq ~ treat)])
```

```{r}
summary(data_treat[, lm(phase_ts ~ treat)])
```

No stat sig relationship between subjective measures and treatment

# Cluster Model

Cluster model. Accounts each individual sleep score per a subject-treat as a cluster.Cluster is per subject, or subject and treat phase?? #TODO

With table in long form, run a simple model and a cluster model for comparison.

```{r}
model_simple = data_long[, lm(sleep_score~treat)]
summary(model_simple)
```

```{r}
# Cluster model, with cluster based on participant_id
model_cluster <- data_long[, lm(sleep_score ~ treat + as.factor(participant_id))]

# We cluster on treat #TODO?? Since it is paired test, we have to differentiate a subjects control and treatment ss as separate clusters
coeftest(model_cluster, vcov = vcovCL(model_cluster, cluster = ss_dt[, participant_id]))
```

# HTE of Assignment Order Subgroups

Difference between Control-Treat, Treat Control? Is there difference in effects based on order of assigned control and treatment?

```{r}
# Non-cluster model, HTEs across treatment order
#order_simple_model <- data_treat[, lm(sleep_score_mean ~ treat + Assignment + (treat * Assignment))]
#summary(order_simple_model)

# Cluster model, HTEs across treatment order
order_cluster_model <- data_long[, lm(sleep_score ~ treat + Assignment + (treat * Assignment) + as.factor(participant_id))]
coeftest(order_cluster_model, vcov = vcovCL(order_cluster_model, cluster = data_long[, participant_id]))
```

Cluster models on treatment order subgroups.

```{r}
# Control then treat
summary(data_long[Assignment == 'ctrl-treat', lm(sleep_score ~ treat + as.factor(participant_id))])
# Treat then control
summary(data_long[Assignment == 'treat-ctrl', lm(sleep_score ~ treat + as.factor(participant_id))])
```

Control-Treatment has positive treatment effect while Treatment-Control exhibited a negative treatment effect. Either effect is statistically significant.

# Effect of Time, or participating in study?

In both treatment order cases, there is an increase in average sleep score from phase 1 to phase 2 (week 1 to week 2) of the study. Could this be an effect of subjects knowing they are being studied (confounder)? This is not a blinded study.

```{r}
# Cluster model, regress sleep scores on week
time_cluster_model <- data_long[, lm(sleep_score ~ week + as.factor(participant_id))]
summary(time_cluster_model)
coeftest(time_cluster_model, vcov = vcovCL(time_cluster_model, cluster = data_long[, participant_id]))
```

Average increase in sleep score by 1.2 from phase 1 for phase 2, regardless of treatment assignment order. Not practically nor statistically significant.

# Other

Other hetergenous treatment effects?

Visual of sleep scores over time

# Using Cluster Individual method for ALL models after SIMPLE

There is alot of data being left on the table. We have to think why we are doing a paired study

1.  It gives us double the amount of data to work with
2.  It controls for inter-subject including inter-device variance. We are leaving alot of this power on the table by not using repeated measures regression.

```{r model with individual as factor}
data$participant_id <- factor(data$participant_id)

model_individual <- lm(sleep_score ~ treat + participant_id, data=data)

robust_individual <- sqrt(diag(vcovHC(model_individual, type="HC1")))

summary(model_individual)
```

```{r model with previous sleep health subjective data}
model_prevsleep <- lm(sleep_score ~ treat + sleep_quality + 
                        trouble_sleep + total_sleep, 
                      data = data)
robust_prevsleep <- sqrt(diag(vcovHC(model_prevsleep, type="HC1")))

summary(model_prevsleep)
```
